Management induced changes of soil organic carbon on global croplands

### 2022-07-12 SOCbudgetPaper/methods.Rmd
update

In a simple first order kinetic approach the steady-state soil organic carbon stocks $SOC^{\mathrm{eq}}$ are given by 
\begin{equation}
SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
(\#eq:inoutflow)
\end{equation}
with $C^{\textrm{in}}$ being the carbon inputs to the soil, $k$ denotes the soil organic carbon decay rate. This equation is valid for all grid cells $i$ and all years $t$. We use the Tier 2 modeling approach for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987]. Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific. 
We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and pastures (referred to as natural vegetation in the following).

@@

Firstly, in order to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we consider land-use change from 1510 onwards. In 1510, we assume all $SOC$ pools to be in natural steady-state, implying that all land-use change prior to that time occurs in 1510. We assume that by 1901, all cropland converted in 1510 has reached its new steady state, so that it is not necessary to explicitly account for even older land conversion. Model inputs for 1901--1930 for climate and natural vegetation litterfall are repeated for 1510--1900 to mimic constant climate conditions for this first initializing period. Similarly, agricultural management data are held constant at the level of 1965 until 1965.  This approach follows others studies looking on effects of land-use change and management [e.g. @schaphoff_lpjml4_2018-1; @herzfeld_soc_2021].

@@

Additionally to the Tier 2 modeling approach [@ogle_cropland_in_ipcc_2019] and the detailed analysis of management data coming with it, $SOC$ changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019]. Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data. Estimates of $F^{\mathrm{SCF}}$ are differentiated by crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these flows. Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig. \@ref(fig:CLIMzone)). The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}}$ by

>>>>

In a simple first order kinetic approach the steady-state soil organic carbon stocks $SOC^{\mathrm{eq}}$ are given by 
\begin{equation}
SOC^{\mathrm{eq}}_{i,t,sub,lu} =\frac{C^{\mathrm{in}}_{i,t,sub,lu}}{k_{i,t,sub,lu}}
(\#eq:inoutflow)
\end{equation}
with $C^{\textrm{in}}$ being the carbon inputs to the soil, $k$ denotes the soil organic carbon decay rate. This equation is valid for all grid cells $i$ and all years $t$. We use the Tier 2 modeling approach for our calculations, which assumes three soil carbon sub-pools $sub$ (active, slow and passive) and interactions between them, following the approach in the Century model [@parton_analysis_1987]. Annual carbon inflow to each sub-pool and annual decay rates of each sub-pool are land-use type $lu$ specific. 
We distinguish two land-use types: cropland and uncropped land under potential natural vegetation as representative for all other land-use types including forestry and grasslands (referred to as natural vegetation in the following). Note that purpose grown annual forage crops are counted as cropland, whereas grazed pastures and mowed meadows (perennials) are assigned to natural vegetation. Carbon flows regarding livestock are split in the sense that only carbon originated from cropland feed sources are considered in this study.

@@

Firstly, in order to account for the impacts of legacy fluxes from land-use changes long before the time horizon of interest, we consider land-use change from 1510 onwards. In 1510, we assume all $SOC$ pools to be in natural steady-state, implying that all land-use change prior to that time occurs in 1510. We assume that by 1901, all cropland converted in 1510 has reached its new steady state, so that it is not necessary to explicitly account for even older land conversion. Model inputs for 1901--1930 for climate and natural vegetation litterfall are repeated for 1510--1900 to mimic constant climate conditions for this first initializing period. Similarly, agricultural management data are held constant at the level of 1965 until 1965. 
We acknowledge that this introduces a bias as agricultural management has changed prior to 1965, but this approach follows others studies looking on effects of land-use change and management [e.g. @schaphoff_lpjml4_2018-1; @herzfeld_soc_2021] and is limited by data availability on harvest statistics (and other management effects).

@@

Additionally to the Tier 2 modeling approach [@ogle_cropland_in_ipcc_2019] and the detailed analysis of management data coming with it, $SOC$ changes can be estimated using the IPCC Tier 1 approach of IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019]. Here, stocks are calculated via stock change factors ($F^{\mathrm{SCF}}$) given by the IPCC for the topsoil (0-30 cm) and based on observational data. Note that IPCC factors are derived under the assumption that there is a linear change between steady states over 20 years. Estimates of $F^{\mathrm{SCF}}$ are differentiated by crops, management and input systems (here summarized under $m$) reflecting different dynamics under changed in- and outflows without explicitly tracking these flows. Moreover, estimates of $F^{\mathrm{SCF}}$ vary for different climatic zones ($c$) specified by the IPCC (see Fig. \@ref(fig:CLIMzone)). The actual $SOC$ stocks are thus calculated based on a given reference stock $SOC^{\mathrm{ref}}$ by
